The Negative Effect of Traffic Noise on House Prices: A Landscape Hedonic Analysis

### 2014-06-23 reports/NoiseHedonicRaR.tex
june 2014 revisions - draft to Tsegaye

\usepackage{alltt}
\usepackage[round]{natbib}
\usepackage[nolists]{endfloat}
\usepackage[width = 5in]{geometry}
\usepackage{caption, amsmath, graphicx, setspace, multirow, color, hyperref, array}

%\renewcommand{\efloatseparator}{\mbox{}}

@@

\title{The Negative Effect of Traffic Noise on House Prices: A Landscape Hedonic Analysis}
\date{}
\author{Aaron Swoboda, Tsegaye Nega and Maxwell Timm}

\doublespacing
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\maketitle
\pagenumbering{roman}
\begin{singlespace}
\begin{abstract}
One consequence of the expanding road network and its associated traffic is increased levels of traffic noise.  While the hedonic literature has consistently shown a negative effect of this phenomenon on the real estate market, research in the United States has often relied on crude measures of traffic noise. Here, we reduce the measurement error of traffic noise exposure through a detailed model of noise propagation over the landscape. Additionally, we estimate the impact on single family home transactions throughout the St.\ Paul, Minnesota, urban area using spatially explicit local regression techniques to allow for spatial non-stationarity in the hedonic function. Contrary to previous work, we find no evidence that the effect of traffic noise varies significantly across the several hundred square miles of our study region, nor by traffic noise levels. We do find significant differences in the impact of traffic noise before and after the economic recession of 2008-09. Our results suggest that an increase in traffic noise of one decibel decreases house prices by an average of 0.19 percent before September 2008 and an average of 0.37 percent after September 2008.
\end{abstract}
\end{singlespace}

@@

Work from Europe suggests that the relationship between traffic noise and house prices varies across housing submarkets \citep{Day2007, MarmolejoDuarteCarlos;GonzalezTamez2009}, by level of traffic noise \citep{Theebe2004a, Brandt2011}, and over time \cite{Wilhelmsson2000}. As such, we construct a flexible hedonic model that allows for a non-stationary relationship between house prices and our explanatory variables over geography and time. Such Locally Weighted Regression (LWR) models have become increasingly common in published work \citep[see][]{MarmolejoDuarteCarlos;GonzalezTamez2009, Carruthers2010, Sunding2010, Nappi-Choulet2011} and have been shown to be more accurate than other common spatial econometric models under common real world circumstances \citep{McMillen2012}. Thus, in addition to using new data that reduces the measurement error of the impact of traffic noise, the spatial and temporal variation in our data and hedonic modeling allow us to present the first estimates in the United States of how the impact of traffic noise varies over space, time, and by noise level.

We find that while the impact of other important control variables in the hedonic function vary significantly over space, the impact of traffic noise does not. Such results are contradictory to recent work by \citet{MarmolejoDuarteCarlos;GonzalezTamez2009} which found evidence of a spatially non-stationary relationship between traffic noise and housing prices using a LWR model in Barcelona, Spain. Our paper also presents the first estimates of the implicit price of traffic noise after the housing market collapse and economic recession of 2008-09. Contrary to work such as \citet{Cho2011b}, which found that environmental amenities mattered \emph{less} after the housing crash, our estimated effect of traffic noise mattered \emph{almost twice as much more} after the recession (0.19 percent reduction in house prices per additional decibel beforehand compared to 0.37 percent per decibel afterwards). We conclude with a discussion of how future research can expand and improve upon this work.

@@

\begin{table}
\caption{Basic Regression Results (dependent variable = $ln(\textrm{Sale Price})$)}\label{tab:globalRegression}

\makebox[\textwidth][c]{%\includegraphics[width = 1.3\textwidth]{../graphs/GlobalRegression}}
% Table generated by Excel2LaTeX from sheet 'Sheet1'
  \centering
  \tabcolsep=0.03cm
    \begin{tabular}{lclcclcclccl}
   & (1)   &       &       & (2)   &       &       & Pre-Crash   &       &       & Post-Crash   &  \\ \hline
\T  Traffic Noise & -2.02E-03 & *** &   & -2.00E-03 & ***   &   & -1.63E-03 & ***  & & -3.08E-03 & *** \\
          & (1.32E-04) &   &   & (1.32E-04) &       &       & (1.38E-04) &       &       & (3.29E-04) &  \\
\T  Intercept & 9.37E+00 & ***   & \hspace{.5cm} & 9.49E+00 & ***   & \hspace{.5cm}   & 8.55E+00 & ***   & \hspace{.5cm}   & 1.10E+01 & *** \\
          & (9.48E-02) &    &   & (9.62E-02) &   &   & (1.11E-01) &       &       & (2.12E-01) &  \\
\T   House Size & 3.38E-04 & ***   &   & 3.39E-04 & ***   &   & 3.35E-04 & ***   &   & 3.50E-04 & *** \\
          & (1.85E-06) &   &   & (1.86E-06) &       &       & (1.99E-06) &       &       & (4.26E-06) &  \\
\T  Lot Size & 1.42E-01 & ***   &    & 2.74E-01 & ***   &   & 2.31E-01 & ***   &      & 2.74E-01 & *** \\
          & (1.14E-02) &     &   & (2.12E-02) &     &      & (2.29E-02) &       &       & (4.99E-02) &  \\
\T  Owner Occupancy & 3.22E-02 & ***   &   & 3.23E-02 & ***   &     & 3.38E-02 & ***  & & 2.61E-02 & *** \\
          & (2.80E-03) &  &   & (2.79E-03) &       &       & (3.24E-03) &       &       & (6.41E-03) &  \\
\T  Year Built & 7.82E-04 & ***   &   & 7.17E-04 & ***   &    & 9.68E-04 & ***   &   & 2.15E-06 &  \\
          & (4.57E-05) &       &       & (4.65E-05) &   &   & (5.06E-05) &    &   & (1.05E-04) &  \\
\T  Median Income & 3.03E-06 & ***  & & 3.31E-06 & ***  &   & 2.86E-06 & ***   &       & 4.06E-06 & *** \\
          & (5.87E-08) &       &       & (5.87E-08) &   &   & (6.56E-08) &       &       & (1.35E-07) &  \\
\T  Elem. Test Scores & 1.88E-03 & *** &   & 1.81E-03 & ***   &   & 3.11E-03 & *** & & 1.21E-03 & *** \\
          & (1.06E-04) &  &   & (1.06E-04) &   &       & (1.75E-04) &       &       & (1.55E-04) &  \\
\T  Distance to CBD & -1.56E-06 & ***  &  & 7.46E-07 & *     &  & -1.57E-07 &   &       & 1.40E-06 &  \\
          & (1.97E-07) &       &       & (3.70E-07) &       &       & (3.95E-07) &   &   & (8.98E-07) &  \\
\T  Distance to Park & -2.38E-07 &   &    & -5.48E-07 &    &   & -6.37E-07 &  &   & -4.01E-07 &  \\
          & (6.82E-07) & &   & (6.83E-07) &       &       & (7.33E-07) &       &       & (1.58E-06) &  \\
\T  Distance to Lake & 2.12E-05 & ***   &   & 2.36E-05 & ***   &   & 2.22E-05 & ***   & & 2.79E-05 & *** \\
          & (1.68E-06) &       &    & (1.71E-06) &   &    & (1.83E-06) & &  & (4.03E-06) &  \\
\T  Distance to Shop & 3.01E-06 & ***   &    & 3.12E-06 & ***   & & 4.16E-06 & ***   &    & 1.88E-06 &  \\
          & (7.67E-07) &       &       & (7.66E-07) &       &       & (8.17E-07) &  & & (1.82E-06) &  \\
\T  Lot Size * Distance to CBD &  &   & & -8.81E-06 & *** & & -6.66E-06 & ***   &       & -9.55E-06 & ** \\
          &   &   &   & (1.19E-06) &       &       & (1.27E-06) &       &       & (2.94E-06) &  \\
\T  County Fixed Effects & yes   &   &   & yes   &       &       & yes   &       &       & yes   &  \\
    Year of Sale Fixed Effects & yes   &   &   & yes   &   &       & yes   &       &       & yes   &  \\
    Month of Sale Fixed Effects & yes   &   &   & yes   &   &       & yes   &       &       & yes   &  \\ \hline
\T  Number of Observations &        42,095  &   &  &    42,095  & & &   31,253  & & &    10,842  &  \\
    R$^2$   & 0.696 &       &       & 0.696 &       &       & 0.714 &       &       & 0.613 &  \\
 \multicolumn{12}{l}{(standard errors) ***$p<.001$; **$p<.01$; *$p<.05$}  \\
    \end{tabular}%
}
\end{table}

Column (1) in Table \ref{tab:globalRegression} displays the results of our basic regression model. All coefficients but one are significantly different from zero at conventional significance levels. The traffic noise coefficient suggests that a one decibel increase traffic noise is on average associated with a 0.2 percent reduction in housing sales prices. Standard urban economic theory \citep{Alonso1964, Mills1967, Muth1969} predicts that the price of land will vary over space within our dataset to account for locational amenities. As such, the other columns include an interaction term between lot size and distance to the nearest central business district to account for spatial variation in the price of land. The significant coefficients on the lot size interaction terms suggest that the value of land varies over space. Other regressions (results available upon request) show similar variation in the implicit prices of house characteristics over space.

Table \ref{tab:globalRegression} also attempts to begin to understand how the economic recession may have influenced the hedonic function in our study area. The third and fourth columns estimate the same regression as (2) but subset the sales before and after September of 2008 (the month that Lehman Brothers filed for bankruptcy, the US Federal government took over Fannie Mae and Freddie Mac, and the American International Group (AIG) narrowly avoided bankruptcy only through a \$85 billion loan from the US Federal government).\footnote{\url{http://www.federalreserve.gov/newsevents/press/other/20080916a.htm}} The coefficients for some of the hedonic variables of interest (noise, schools, land, for instance) appear different across the two time periods.

The results in Table \ref{tab:globalRegression} and previous work \citep[such as][]{Day2007, MarmolejoDuarteCarlos;GonzalezTamez2009} suggests that the relationship between prices and many other hedonic characteristics may vary over space. While methods exist for parameterizing this variation \citep[such as spatial expansion as suggested by][]{Casetti1972}, we have no a priori knowledge of how to parameterize the variation over space and time. As such, we turn to a semi-parametric form of hedonic regression, a flexible modeling approach which lets the data reveal how relationships vary, rather than specifying them beforehand. Such a modeling choice will allow us to test the hypotheses of spatial and temporal stationarity in the regression model.

\section{Locally Weighted Regression Model}
Locally Weighted Regression (LWR) techniques (also known as Geographically Weighted Regression) are described in detail by \citet{Cleveland1988}, \citet{Brunsdon1998b}, \citet{Fotheringham2002}, and others. It is a weighted least squares methodology in which regression coefficients are estimated over space as a function of the local data as described in Equation~\eqref{eq:LWRbeta},
\begin{equation}\label{eq:LWRbeta}
\hat{\beta}_i = (X'W_iX)^{-1}X'W_iY,
\end{equation}
where X is a $n \times m$ matrix of independent variables, $W_i$ is the $n \times n$ weights matrix, and Y is the $n \times 1$ vector of dependent variable values. The weights matrix, $W_i$ is a diagonal matrix where element $w_{jj}$ denotes the weight that the $j^{th}$ data point will receive in the regression coefficients estimated at location $i$ in the dataset. We employ a bi-square weights function and a k-nearest neighbor bandwidth approach as described in equation~\eqref{eq:weights}, 
\begin{equation}\label{eq:weights}
w_{jj}=\left[1-\left(\frac{d_{ij}}{d_{k}}\right)^2 \right]^2 \textrm{ if  }d_{ij}<d_{ik}\textrm{, otherwise = 0},
\end{equation}
where $d_{ij}$ denotes the distance between observations $i$ and $j$, and $d_{ik}$ is the distance from observation $i$ to the $k^{th}$ nearest observation. This function assigns weights close to 1 for data points near observation $i$, weights positive but closer to zero for observations farther away, and zero for all $n-k$ observations farther away than the $k^{th}$ nearest observation. We estimate LWR coefficients using bandwidths ranging from as small as $k=50$ observations and as large as 10,000 observations. We choose the LWR bandwidth my minimizing the Generalized Cross Validation score as detailed in equation~\eqref{eq:GCV},
\begin{equation}\label{eq:GCV}
n*\sum_{i=1}^{n}\frac{(y_i-\hat{y}_i)^2}{(n-v_1)^2}, 
\end{equation} 
where $y_i$ is the dependent variable value, $\hat{y}_i$ is the predicted dependent variable value for observation $i$, and $v_1$ is the ``effective number of model parameters.''
$v_1=$tr(\textbf{S}), where the matrix \textbf{S} is the ``hat matrix'' which maps $y$ onto $\hat{y}$,
                   \begin{equation*}
                   \hat{y}=\textbf{S}y,
                   \end{equation*}
                   and each row of \textbf{S}, $r_i$ is given by:
                     \begin{equation*}
                   r_i=X_i(X'W_iX)^{-1}X'W_i.
                   \end{equation*}
The GCV score is a convenient model selection metric that rewards models that provide a good fit to the data, while penalizing models with a greater number of model parameters \citep{Loader1999, McMillen2010}. Model selection via this strategy has been shown to discern whether spatially varying relationships exist, accurately estimate spatially varying coefficients, even outperform other spatial econometric techniques, and do so without wasting degrees of freedom \citep{Paez2011, McMillen2010, McMillen2012}.

%\section{Results}
Similar to the basic econometric model described in Section \ref{basicModel}, the LWR model estimates the logged sales price of a house as a function of structural variables, location and time. In order to account for changing market conditions in our data, when estimating LWR coefficients we use only houses sold within the past 12 months. Thus, the coefficient estimates at a particular house are estimated using data from other sales nearby in both time and geography.

\subsection{Improved Model Performance}
The use of the Locally Weighted Regression model significantly increases model performance. Figure \ref{fig:GCVmodel} displays the Generalized Cross Validation score for three different model specifications with varying number of observations included in the regression bandwidth. Two important conclusions can be drawn from the results. First, local analysis (bandwidths of a couple hundred observations) perform significantly better than global models, and second, conditional on the bandwidth, performance improves with the inclusion of more locational variables. The minimum GCV score is obtained using a bandwidth of 200 nearest houses and a model including structural, locational, and city fixed-effects.

We estimate three different models. Model (1) uses only the basic structural characteristics as the explanatory variables (size of the house in square feet, size of the lot in acres, a categorical variable denoting the architectural style of the house, and whether or not the house is owner-occupied). Model (2) adds locational variables to the previously described model (test scores at the local elementary school, census tract median income, and distances to the central business district, nearest park, nearest shopping center, and nearest lake). The third model adds city fixed effects. Generally speaking, the third model performs slightly better than the second model, which performs substantially better than the first model. 
\begin{figure}
  \makebox[\textwidth][c]{\includegraphics[width = 1.2\textwidth]{../graphs/GCVbyModel}}
\caption{This figure shows the relationship between bandwidth size and the GCV score for three different Locally Weighted Regression (LWR) models. For comparison, the GCV scores for each model when estimated at a global scale are also shown. Note that the LWR models all have significantly smaller GCV scores than the global models. The minimum GCV score is obtained by LWR Model (3) at a bandwidth of 200 nearest houses.}\label{fig:GCVmodel}
\end{figure}

We also estimate each of the three models at varying levels of local analysis, with bandwidths ranging from the nearest 50 observations to the nearest 10,000. Models with bandwidths of only a couple hundred observations have substantially smaller GCV scores, suggesting that, conditional on control variables included in the model, model performance still improves by performing the analysis at the local level. It is also worth noting that the LWR models have significantly lower GCV scores for a range of bandwidths when compared to the global models. These results suggest that it is important to model our data in a spatially explicit manner.

Table \ref{tab:mainresults} displays a summary of the results of LWR Model (3) and a similar regression model run without allowing for spatial variation in the regression coefficients. Again, note that the LWR model yields a substantially smaller GCV score (0.027 vs.\ 0.041) signifying a substantial improvement in model performance. In addition to the significant reduction in the GCV score obtained with the use of LWR, we also see a substantial reduction in the degree of spatial autocorrelation in the model residuals. We calculate Moran's I statistic to be 0.199 for the global model while our preferred LWR specification yields 0.012, a more than ten-fold reduction.

% latex table generated in R 2.15.2 by xtable 1.7-0 package
% Mon Jun 24 16:36:58 2013

>>>>

\usepackage{alltt}

\usepackage[round]{natbib}
\usepackage{dcolumn}
\usepackage[nolists]{endfloat}
\usepackage[width = 5in]{geometry}
\usepackage{caption, amsmath, graphicx, setspace, multirow, color, hyperref, array}
\usepackage[modulo]{lineno}
%\renewcommand{\efloatseparator}{\mbox{}}

@@

\title{The Negative Effect of Traffic Noise on House Prices: A Landscape Hedonic Analysis}
\date{}
\author{}%Aaron Swoboda, Tsegaye Nega and Maxwell Timm}

\doublespacing
\begin{document}
\maketitle
\pagenumbering{roman}
\begin{singlespace}
\begin{abstract}
One consequence of the expanding road network and its associated traffic is increased levels of traffic noise.  While the hedonic literature has consistently shown a negative effect of this phenomenon on the real estate market, research in the United States has often relied on crude measures of traffic noise. Here, we reduce the measurement error of traffic noise exposure through a detailed model of noise propagation over the landscape. Additionally, we estimate the impact on single family home transactions throughout the St.\ Paul, Minnesota, urban area using spatially explicit local regression techniques to allow for spatial non-stationarity in the hedonic function. 
\end{abstract}
\end{singlespace}

@@

We construct a flexible hedonic model that allows for a non-stationary relationship between house prices and our explanatory variables over geography and time. Such Locally Weighted Regression (LWR) models have become increasingly common in published work \citep[see][]{MarmolejoDuarteCarlos;GonzalezTamez2009, Carruthers2010, Sunding2010, Nappi-Choulet2011} and have been shown to be more accurate than other common spatial econometric models under common real world circumstances \citep{McMillen2012}. Thus, in addition to using new data that reduces the measurement error of the impact of traffic noise, the spatial and temporal variation in our data and hedonic modeling allow us to present the first estimates in the United States of how the marginal effect of traffic noise varies over space and time.

We find strong evidence to support hedonic modeling choices that allow for spatial and temporal heterogeneity. Our paper also presents the first estimates of the implicit price of traffic noise after the housing market collapse and economic recession of 2008-09. Contrary to work such as \citet{Cho2011b}, which found that environmental amenities mattered \emph{less} after the housing crash, our estimated noise coefficients roughly double in magnitude after the recession compared to before. We conclude with a discussion of how future research can expand and improve upon this work.

@@

\begin{table}
\caption{Basic Global Regression Results}\label{tab:globalRegression}
\makebox[\textwidth][c]{
\scriptsize
  \centering
\begin{tabular}{@{\extracolsep{-1pt}}lD{.}{.}{-4} D{.}{.}{-4} D{.}{.}{-4} D{.}{.}{-4} } 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable: ln Sale Price}} \\ 
\cline{2-5} \\[-1.8ex] 
& \multicolumn{1}{r}{Model (A)} & \multicolumn{1}{r}{Model (B)} & \multicolumn{1}{r}{Model (C)} & \multicolumn{1}{r}{Model (D)}\\ 
& \multicolumn{1}{r}{All Years} & \multicolumn{1}{r}{Year=2006} & \multicolumn{1}{r}{Year=2008} & \multicolumn{1}{r}{Year=2010}\\
\hline \\[-1.8ex] 
Noise (dB) & -0.0027^{***} & -0.0020^{***} & -0.0033^{***} & -0.0040^{***} \\ 
  & (0.0001) & (0.0003) & (0.0005) & (0.0006) \\ 
  & & & & \\ 
 House Size (1,000s ft$^2$) & 0.2914^{***} & 0.2847^{***} & 0.2810^{***} & 0.3066^{***} \\ 
  & (0.0022) & (0.0043) & (0.0067) & (0.0084) \\ 
  & & & & \\ 
 Lot Size (acres) & 0.2493^{***} & 0.2719^{***} & 0.2799^{***} & 0.2135^{***} \\ 
  & (0.0118) & (0.0221) & (0.0359) & (0.0447) \\ 
  & & & & \\ 
 Owner Occupancy Dummy & 0.0222^{***} & -0.0027 & 0.0478^{***} & 0.0298^{***} \\ 
  & (0.0026) & (0.0050) & (0.0085) & (0.0097) \\ 
  & & & & \\ 
 Year House Built & 0.0020^{***} & 0.0025^{***} & 0.0018^{***} & 0.0011^{***} \\ 
  & (0.0001) & (0.0001) & (0.0002) & (0.0002) \\ 
  & & & & \\ 
 Percent Under 18 & -0.0001 & 0.00002 & -0.0002 & -0.0006 \\ 
  & (0.0001) & (0.0002) & (0.0004) & (0.0005) \\ 
  & & & & \\ 
 Percent White & 0.0037^{***} & 0.0031^{***} & 0.0040^{***} & 0.0040^{***} \\ 
  & (0.0001) & (0.0001) & (0.0002) & (0.0003) \\ 
  & & & & \\ 
 Median Income (\$1,000s) & 0.0023^{***} & 0.0021^{***} & 0.0023^{***} & 0.0029^{***} \\ 
  & (0.0001) & (0.0001) & (0.0002) & (0.0003) \\ 
  & & & & \\ 
 Elementary Test Scores & 0.0009^{***} & 0.0023^{***} & 0.0040^{***} & 0.0001 \\ 
  & (0.0001) & (0.0003) & (0.0005) & (0.0002) \\ 
  & & & & \\ 
 Dist to CBD (km) & 0.0078^{***} & 0.0083^{***} & 0.0053^{***} & 0.0101^{***} \\ 
  & (0.0005) & (0.0010) & (0.0016) & (0.0020) \\ 
  & & & & \\ 
 Dist to Lake (km) & 0.0230^{***} & 0.0246^{***} & 0.0021 & 0.0343^{***} \\ 
  & (0.0017) & (0.0032) & (0.0054) & (0.0067) \\ 
  & & & & \\ 
 Dist to Park (km) & 0.0014 & -0.0027 & 0.0099^{***} & 0.0064^{*} \\ 
  & (0.0009) & (0.0017) & (0.0027) & (0.0034) \\ 
  & & & & \\ 
 Dist to Shop (km) & -0.0019^{*} & -0.0020 & -0.0063^{*} & -0.0020 \\ 
  & (0.0010) & (0.0019) & (0.0033) & (0.0041) \\ 
  & & & & \\ 
\hline \\[-1.8ex] 
Home Style Fixed Effects & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} \\ 
Month*Year Fixed Effects & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} \\ 
City Fixed Effects & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{Yes} \\ 
\hline \\[-1.8ex] 
Observations & \multicolumn{1}{c}{42,083} & \multicolumn{1}{c}{8,885} & \multicolumn{1}{c}{5,498} & \multicolumn{1}{c}{4,200} \\ 
R$^{2}$ & \multicolumn{1}{c}{0.742} & \multicolumn{1}{c}{0.785} & \multicolumn{1}{c}{0.691} & \multicolumn{1}{c}{0.668} \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
}
\end{table}

Model (A) in Table \ref{tab:globalRegression} displays the results of a basic Ordinary Least Squares regression model. In addition to the important structural and neighborhood variables, we also include fixed effects for each city and month of sale in the dataset to allow for month-to-month shocks to prices. The traffic noise coefficient suggests houses in our data that are similar in all respects but a one decibel difference in noise will, on average, have sales prices roughly 0.27 percent lower at the noisier location.

Further exploratory analysis suggests the assumption of temporal and geographic stationarity of the hedonic price function is overly restrictive and needs to be relaxed. We first tested for temporal changes in the coefficients for house size, lot size, and noise by creating interaction terms with these variables and the month*year year fixed effects. In all three cases the improved model fit was significantly more than expected due to random chance.\footnote{The results of specific hypothesis tests are available upon request.} Columns (B) - (D) of Table \ref{tab:globalRegression} show the results of restimating Model (A) using only sales within three separate years of data. Subsequent analysis also suggests that many of the important regression coefficients vary across the different cities within our data. We are able to reject the null hypothesis that the interaction terms between cities and important variables (house size, lot size, noise, etc.) are zero, even after limiting the analysis to only houses sold within a given year (for instance, by adding the interaction terms to model specifications (B) - (D)).\footnote{Results of specific F-tests are available upon request.}

These results should be taken with appropriate skepticism, as the models are simple Ordinary Least Squares regressions and the spatial and temporal nature of the data likely violate assumptions of the Classical Linear Regression Model. However, the initial findings seem to suggest that the relationship between house prices and many other hedonic characteristics may vary over space and time. While methods exist for parameterizing this variation \citep[such as spatial expansion as suggested by][]{Casetti1972}, we have no a priori knowledge of how to parameterize the variation. As such, we turn to a semi-parametric form of hedonic regression, a flexible modeling approach which lets the data reveal how relationships vary, rather than specifying them beforehand. Such a modeling choice will allow us to perform additional tests of the hypotheses of spatial and temporal stationarity in the regression model.

\section{Locally Weighted Regression Model}
Locally Weighted Regression (LWR) techniques (also known as Geographically Weighted Regression) are described in detail by \citet{Cleveland1988}, \citet{Brunsdon1998b}, \citet{Fotheringham2002}, and others. It is a weighted least squares methodology in which regression coefficients are estimated over space as a function of the local data as described in Equation~\eqref{eq:LWRbeta},
\begin{equation}\label{eq:LWRbeta}
\hat{\beta}_i = (X'W_iX)^{-1}X'W_iY,
\end{equation}
where X is a $n \times m$ matrix of independent variables, $W_i$ is the $n \times n$ weights matrix, and Y is the $n \times 1$ vector of dependent variable values. The weights matrix, $W_i$ is a diagonal matrix where element $w_{jj}$ denotes the weight that the $j^{th}$ data point will receive in the regression coefficients estimated at location $i$ in the dataset. We employ a bi-square weights function and a k-nearest neighbor bandwidth approach as described in equation~\eqref{eq:weights}, 
\begin{equation}\label{eq:weights}
w_{jj}=\left[1-\left(\frac{d_{ij}}{d_{k}}\right)^2 \right]^2 \textrm{ if  }d_{ij}<d_{ik}\textrm{, otherwise = 0},
\end{equation}
where $d_{ij}$ denotes the distance between observations $i$ and $j$, and $d_{ik}$ is the distance from observation $i$ to the $k^{th}$ nearest observation. This function assigns weights close to 1 for data points near observation $i$, weights positive but closer to zero for observations farther away, and zero for all $n-k$ observations farther away than the $k^{th}$ nearest observation. We estimate LWR coefficients using bandwidths ranging from as small as $k=50$ observations and as large as 10,000 observations. We choose the LWR bandwidth my minimizing the Generalized Cross Validation score as detailed in equation~\eqref{eq:GCV},
\begin{equation}\label{eq:GCV}
n*\sum_{i=1}^{n}\frac{(y_i-\hat{y}_i)^2}{(n-v_1)^2}, 
\end{equation} 
where $y_i$ is the dependent variable value, $\hat{y}_i$ is the predicted dependent variable value for observation $i$, and $v_1$ is the ``effective number of model parameters.'' The value  
$v_1=$tr(\textbf{S}), where the matrix \textbf{S} is the ``hat matrix'' which maps $y$ onto $\hat{y}$,
                   \begin{equation*}
                   \hat{y}=\textbf{S}y,
                   \end{equation*}
                   and each row of \textbf{S}, $r_i$ is given by:
                     \begin{equation*}
                   r_i=X_i(X'W_iX)^{-1}X'W_i.
                   \end{equation*}
The GCV score is a convenient model selection metric that rewards models that provide a good fit to the data, while penalizing models with a greater number of model parameters \citep{Loader1999, McMillen2010}. Model selection via this strategy has been shown to discern whether spatially varying relationships exist, accurately estimate spatially varying coefficients, even outperform other spatial econometric techniques, and do so without wasting degrees of freedom \citep{Paez2011, McMillen2010, McMillen2012}.

%\section{Results}
Similar to the basic econometric model described in Section \ref{basicModel}, the LWR model estimates the logged sales price of a house as a function of structural and neighborhood variables, location and time. In order to account for changing market conditions in our data as manifested in regression coefficients that vary over time, when estimating LWR coefficients we use only houses sold within the past 12 months and also include month fixed effects to account for seasonality. Thus, the coefficient estimates at a particular house are estimated using data from other sales nearby in both time and geography.

\subsection{Improved Model Performance}
The use of the Locally Weighted Regression model significantly increases model performance as measured by the generalized cross validation score. Figure \ref{fig:GCVmodel} displays the Generalized Cross Validation score for three different model specifications estimated at the ``global'' level as well as with varying numbers of observations included in the regression bandwidth. We estimate three different hedonic function specifications with different levels of additional spatial control variables. Model (1) includes the noise variable, the basic structural characteristics (size of the house in square feet, size of the lot in acres, a categorical variable denoting the architectural style of the house, and whether or not the house is owner-occupied) plus month of sale*year fixed effects. Model (2) includes all of the str adds neighborhood and locational variables to the previously described model (test scores at the local elementary school, census tract median income, percent of the census block population that is white, percentage of the census block that is under age 18 and distances to the central business district, nearest park, nearest shopping center, and nearest lake). The third model adds city fixed effects to the second model. Thus Model (3) explicitly includes both common important characteristics of neighborhood quality while also allowing for spatial heterogeneity in the regression coefficients, and even city-level fixed effects to control for omitted variables that vary at the municipal level.

Two important conclusions can be drawn from these results. Model (3) consistently outperforms the other two models-- slightly better than Model (2) and considerably better than Model (1). Thus, even conditional on having modeled the data with local regression, the inclusion of more neighborhood-level variables substantially improves the model fit. Second, local analyses perform significantly better than global models. The minimum GCV score is obtained using a bandwidth of 500 nearest houses and a model including structural, neighborhood, and locational variables in addition to time and city fixed-effects.

\begin{figure}
  \makebox[\textwidth][c]{\includegraphics[width = 1.2\textwidth]{../graphs/GCVbyModel}}
\caption{This figure shows the relationship between bandwidth size and the GCV score for three different Locally Weighted Regression (LWR) models. For comparison, the GCV scores for each model when estimated at a global scale are also shown. Note that the LWR models all have significantly smaller GCV scores than the global models. The minimum GCV score is obtained by LWR Model (3) at a bandwidth of 500 nearest houses.}\label{fig:GCVmodel}
\end{figure}

%Table \ref{tab:LWR} summarizes the results of LWR Model (3) at three different bandwidth sizes plus the same model estimated at the ``global'' level. run without allowing for spatial variation in the regression coefficients. Again, note that the LWR model yields a substantially smaller GCV score (0.027 vs.\ 0.038) signifying a substantial improvement in model performance. In addition to the significant reduction in the GCV score obtained with the use of LWR, we also see a substantial reduction in the degree of spatial autocorrelation in the model residuals. We calculate Moran's I statistic to be 0.199 for the global model while our preferred LWR specification yields 0.012, a more than ten-fold reduction.
